A Large-Scale Study of the Use of Eval in R

### 2020-08-23 inst/analysis/corpus-analysis.Rmd
First paragraph in 5

---
title: "Corpus Analysis"
output: 
  rmdformats::readthedown:
        code_folding: hide
        lightbox: true
        gallery: false
        df_print: paged
        toc_depth: 3
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "html") })
---

>>>>

---
title: "Corpus Analysis"
output: html_document
params:
  base_dir: ../../../
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
for (x in c(
  "dplyr", 
  "DT",
  "fs",
  "fst",
  "ggplot2",
  "ggthemes",
  "lubridate",
  "readr",
  "purrr",
  "stringr",
  "tidyr"
  )) {
  suppressPackageStartupMessages(library(x, character.only=TRUE))
}

knitr::opts_chunk$set(
  echo = TRUE,
  fig.align = "center",
  fig.retina = 2,
  fig.width = 10
)

# TODO: move into the package
source("inc/paths.R")
source("inc/setup.R")
source("inc/latextags.R")

options(repos="https://cran.r-project.org")
theme_set(theme_minimal())
create_tags(path(TAGS_DIR, "corpus.tex"), prefix="Corpus", default=TRUE)
```

## Load data

```{r}
corpus <- as_tibble(read_fst(CORPUS_DETAILS_FILE))
```

## Corpus Overview

### Table

```{r summary}
overview_table(
  r("packages",             corpus),
  r("runnable code",        sum(corpus$runnable_code, na.rm=T)),
  r("examples code",        sum(corpus$runnable_code_examples, na.rm=T)),
  r("tests code",           sum(corpus$runnable_code_tests, na.rm=T)),
  r("vignettes code",       sum(corpus$runnable_code_vignettes , na.rm=T)),
  r("mean expr coverage",   percent(mean(corpus$coverage_expr))),
  r("median expr coverage", percent(median(corpus$coverage_expr))),
  r("with R code",          filter(corpus, package_r_code > 0)),
  r("R code",               sum(corpus$package_r_code, na.rm=T)),
  r("with native code",     filter(corpus, package_native_code > 0)),
  r("native code",          sum(corpus$package_native_code, na.rm=T)),
  r("functions",            sum(corpus$funs_public+corpus$funs_private, na.rm=T)),
  r("public functions",     sum(corpus$funs_public, na.rm=T)),
  r("private functions",    sum(corpus$funs_private, na.rm=T)),
  r("ranked in cranlogs",   filter(corpus, !is.na(rank))),
  r("mean revdes",          mean(corpus$revdeps)),
  r("median revdes",        median(corpus$revdeps)),
  r("runtime",              as.duration(sum(corpus$runnable_time))),
  r("mean runtime",         as.duration(mean(corpus$runnable_time))),
  r("median runtime",       as.duration(median(corpus$runnable_time))),
  r("function with eval",   sum(corpus$funs_with_eval, na.rm=T))
)
```

### CRAN

```{r}
overview_table(
  r("all cran", dim(available.packages())[1])
)
```

### Plots

```{r plot corpus}
coverage_mean <- mean(corpus$coverage_expr)
coverage_median <- median(corpus$coverage_expr)
revdeps_mean <- mean(corpus$revdeps)
revdeps_median <- median(corpus$revdeps)

corpus %>%
  mutate(
    outlier=package_code > PACKAGE_SIZE_OUTLIER,
    label=ifelse(rank <= 25 | outlier, package, NA)  #if_else(outlier, str_glue("{package}"), as.character(NA))
  ) %>%
  ggplot(
    aes(
      x=coverage_expr, 
      y=revdeps,
      label=label,
      size=package_code
    )
  ) +
  geom_point(alpha=.6) +
  geom_text(size=2, check_overlap = T, vjust=1.5, na.rm = TRUE) +

  geom_hline(aes(yintercept=revdeps_mean), linetype="dashed", color="red", size=0.2) +
  geom_vline(aes(xintercept=coverage_mean), linetype="dashed", color="red", size=0.2) +

  scale_x_continuous(labels = scales::percent) + 
  scale_y_log10(labels = scales::comma) +
  scale_size(range = c(.1, 3), labels = scales::comma) +
  theme(
    legend.position=c(0.2, 0.85),
    legend.box="horizontal",
    legend.box.background = element_rect(fill="white", size=0.1)
  ) +
  labs(
    x="Code coverage", 
    y="Number of reverse dependencies (log)", 
    size="Code size (R + native)",
    linetype=""
  ) +
  guides(
    size=guide_legend(order=1),
    color=guide_legend(order=2),
    linetype=F
  )

ggsave(CORPUS_PLOT)
```

### List of corpus packages

```{r list of packages, eval=FALSE}
datatable(
  corpus, 
  filter="top", 
  options=list(
    pageLength=50,
    autoWidth=TRUE
  )
)
```

## Kaggle

```{r, eval=FALSE}
parallel_log <- 
  read_csv(PARALLEL_LOG_FILE)

parallel_log %>%
  mutate(
    output=show_url(path(V1, "run.R.out"))
  ) %>%
  my_datatable(escape=FALSE)
```

### 2020-08-23 inst/paper/main.tex
First paragraph in 5

\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{listings}
\newcommand{\missingNumber}{\textcolor{red}{XX}\xspace}
\newcommand{\missingPercentage}{\textcolor{red}{XX\%}\xspace}
\newcommand{\missingTable}{\textcolor{red}{XXTable}\xspace}
\newcommand{\missingGraph}{\textcolor{red}{XXGraph}\xspace}
\usepackage[usenames,dvipsnames]{xcolor} 
\usepackage{xspace}

\input{counts.tex}

@@

@@

>>>>

\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{listings}
\newcommand{\missingNumber}{\textcolor{red}{XX}\xspace}
\newcommand{\missingPercentage}{\textcolor{red}{XX\%}\xspace}
\newcommand{\missingTable}{\textcolor{red}{XXTable}\xspace}
\newcommand{\missingGraph}{\textcolor{red}{XXGraph}\xspace}
\usepackage[usenames,dvipsnames]{xcolor} 
\usepackage{xspace}
\usepackage{url}

\input{counts.tex}
\input{tag/corpus.tex}

@@

\newcommand{\eg}{\emph{e.g.},\xspace}
\newcommand{\ie}{\emph{i.e.},\xspace}
\newcommand{\cf}{\emph{cf.}\xspace}

@@

We have selected \CorpusPackages packages consisting of \CorpusRCodeRnd lines of
R code and \CorpusNativeCodeRnd lines of native code (C/C++/Fortran).
Figure~\ref{fig:corpus} shows these packages, the size of the dots reflects the
project's size in lines of code including both R and native code\footnote{Lines
  of source code reported excludes comments and blank lines, counted by
  \emph{cloc}, \cf \url{https://github.com/AlDanial/cloc}}, the x-axis indicates
the expression code coverage in percents and the y-axis gives the number of
reverse dependencies in log scale. Dotted lines indicate means. Large and
popular packages are annotated.

\begin{figure*}[!tb]\centering\includegraphics[width=.9\linewidth]
  {corpus.pdf}\caption{Corpus}\label{fig:corpus}
\end{figure*}

These packages come from the Comprehensive R Archive Network
(CRAN\footnote{\url{http://cran.r-project.org}}), the largest repository of R
code with over \CorpusAllCranRnd packages\footnote{CRAN receives about 6 new
  package submissions a day~\cite{Ligges2017}}. Unlike other open code
repositories such as GitHub, CRAN is a curated repository. Each submitted
package must abide to a number of well-formedness rules that are automatically
checked asserting certain quality. Most relevant for this work is that all of
the runnable code (including code snippets from examples and vignettes) is
tested and only a successfully running package is admitted in the archive.

% FILIP: WIP
% We have downloaded and installed all available CRAN packages. Out of the
% \AllCranRnd packages, we managed to install \AllLoadableRnd. The main reason for
% this is that \rdt is based on GNU-R 3.5.0 and some of the packages are no longer
% compatible with this version. Some packages also require extra native
% dependencies which were not present on our servers. We defined two criteria for
% including a package into the corpus:
% \begin{inparaenum}[(1)]
%   \item have a runnable code that covers a significant part of the package
%   source code from which type signatures could be inferred, and
%   \item have some reverse dependencies that will allows us to evaluate the
%   inferred types using the runnable code from these dependencies.
% \end{inparaenum}
% The concrete thresholds used were: at least \ThresholdCodeCoverage of expression
% coverage and at minimum \ThresholdRevdeps reverse dependencies. The code
% coverage was computed for each package using
% \covr\footnote{\url{https://github.com/r-lib/covr}}, the R code coverage R tool.
% The reverse package dependencies were extracted from the package metadata using
% builtin function.

% The \CorpusLoadable selected packages contain \CorpusRunnableCodeRnd lines of
% runnable code in examples (\CorpusExamplesCodeRnd), tests (\CorpusTestsCodeRnd)
% and vignettes (\CorpusVignettesCodeRnd). Running this code results on average in
% \CorpusMeanExprCoverage package code coverage (CRAN average is
% \AllMeanExprCoverage). Together, there is \CorpusRevdesRnd (on average
% \CorpusMeanRevdes, median \CorpusMedianRevdes; CRAN average is \AllMeanRevdes,
% median \AllMedianRevdes) reverse dependencies with \CorpusRevdepRunnableCodeRnd
% runnable lines of code resulting in \CorpusRevdepMeanCodeCoverage coverage (on
% average) of the corpus packages.
% Together there are \CorpusFunctionsRnd defined R functions
% (\CorpusPublicFunctionsRnd are from packages' public API).
% \CorpusSThreeFunctionsRnd are S3 functions. Packages in the corpus define
% \CorpusSThreeClasses S3 classes.
% The code coverage was computed for each package using
% \covr\footnote{\url{https://github.com/r-lib/covr}}, the R code coverage R
% tool.  The reverse package dependencies were extracted from the package
% metadata using builtin function.
